---
title: "Lesson01-PCA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root_path <- "C:/Git/RCourse_ML"
exerices_path <- file.path(root_path, "Exercises_and_Solutions")
```

## Lesson 01 - PCA

```{r}
lesson_path <- file.path(exerices_path, "Lesson01_PCA")
```


load CSV
```{r}
bcw <- read.csv(file.path(lesson_path, "breastCancer_Wisconsin.csv"), stringsAsFactors = FALSE)
```


There are `rnrow(bcw)-1` variables, of each 32 observations

# run PCA

remove the id colum (incremental ids) and the diagnosis (categorial) column
```{r}
pca_dat <- bcw[, 3:ncol(bcw)]
```

```{r}
pca_1 <- prcomp(pca_dat, scale = TRUE)
pca_2 <- prcomp(pca_dat, scale = FALSE)
```

```{r}
summary(pca_1)
```
PC1 explains `r summary(pca_1)$importance[2, 'PC1']` of the variance, the first four PCs explains `r summary(pca_1)$importance[3, 'PC4']`.

The variable contributing most to PC1 is (in decreasing order):
```{r}
t_val <- "PC1"
head(pca_1$rotation[order(abs(pca_1$rotation[, t_val]), decreasing = TRUE), t_val])

```

coordinate fo the first obs for the first three PCs are:
```{r}
pca_1$x[1, 1:3]
```


## Visualise biplot


```{r}
plot(PC2 ~PC1, data = pca_1$x, col = as.factor(bcw$diagnosis))
legend("bottomleft", c("benign", "malignat"), pch = c(19,19), col = 1:2, cex = 1.7)
```


compare PCA without scaling

```{r}
plot(PC2 ~ PC1, data = pca_2$x, col = as.factor(bcw$diagnosis))
legend("bottomleft", c("benign", "malignat"), pch = c(19,19), col = 1:2, cex = 1.7)
```

```{r}
library(ggplot2)
library(ggfortify)
bcw$col <- ifelse(bcw$diagnosis == "B", 1, 2)
autoplot(pca_1, col = as.factor(bcw$col))
```



## Scree plot
```{r}

screeplot(pca_1, npcs = 10, type = 'l')

```

