---
title: "Lesson03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root_path <- "C:/Git/RCourse_ML"
exerices_path <- file.path(root_path, "Exercises_and_Solutions")
```

```{r}
library(MASS) # to load dataset
library(class)

```


## Lesson 03 - KNN

```{r}
lesson_path <- file.path(exerices_path, "Lesson03_KNN")
```


### exercies 1
```{r}
str(crabs)
table(crabs$sex, crabs$sp)
```
keep predictors
```{r}
knn_dat <- crabs[4:8]
```

#### fit a KNN with 3 neighbours for SPECIES
```{r}
knn.pred <- knn(train = knn_dat, # use the iris data as training data
  test = knn_dat, # knn always needs a specified test set
  cl = crabs$sp, # correct labels of training set
  k = 3) # parameter k

```


Confusion matrix:
```{r}
confT <- table(knn.pred, crabs$sp)
confT
```
Training error:
```{r}
missT <- confT
diag(missT) <- 0 # turn diagonal values to zeros
missCount <- sum(missT) # count the miss-classifications (5)
trainErr <- missCount/nrow(crabs)
trainErr
```
#### fit a KNN with 3 neighbours for SEX

```{r}
knn.pred.sex <- knn(train = knn_dat, # use the iris data as training data
  test = knn_dat, # knn always needs a specified test set
  cl = crabs$sex, # correct labels of training set
  k = 3) # parameter k

```
Confusion matrix:
```{r}
confT <- table(knn.pred.sex, crabs$sex)
confT
```
Training error:
```{r}
missT <- confT
diag(missT) <- 0 # turn diagonal values to zeros
missCount <- sum(missT) # count the miss-classifications (5)
trainErr <- missCount/nrow(crabs)
trainErr
```


#### fit a KNN with 3 neighbours for SEXxSpecies

```{r}
crabs$sp_sex <- interaction(crabs$sp, crabs$sex)
table(crabs$sp_sex)
```


```{r}
knn.pred.spsex <- knn(train = knn_dat, # use the iris data as training data
  test = knn_dat, # knn always needs a specified test set
  cl = crabs$sp_sex, # correct labels of training set
  k = 3) # parameter k

```
Confusion matrix:
```{r}
confT <- table(knn.pred.spsex, crabs$sp_sex)
confT
```
Training error:
```{r}
missT <- confT
diag(missT) <- 0 # turn diagonal values to zeros
missCount <- sum(missT) # count the miss-classifications (5)
trainErr <- missCount/nrow(crabs)
trainErr
```

### exercies 2
load CSV
```{r}
taxi <- read.csv(file.path(lesson_path, "taxi.csv"))
str(taxi)
table(taxi$col)
```
make colour variabel as factor
```{r}
taxi$col_f <- as.factor(taxi$col)
```

run KNN using a 1000 test samples, and 4 neighbours
```{r}
set.seed(111)
N <- 1000
index <- sample(1:nrow(taxi), N, replace = FALSE)
taxi_train <- taxi[-test_index, ]
taxi_test <- taxi[test_index, ]
```


```{r}
knn_pred_tr <- knn(train = taxi_train[, 1:6],
                 test = taxi_train[, 1:6],
                 cl = taxi_train[, "col"],
                 k = 4
                 )
```

Confusion matrix:
```{r}
confT_tr <- table(knn_taxi1, taxi[test_index, "col"])
confT_tr
```
Training error:
```{r}
missT <- confT_tr
diag(missT) <- 0 # turn diagonal values to zeros
missCount <- sum(missT) # count the miss-classifications (5)
trainErr_tr <- missCount/nrow(taxi_train)
trainErr_tr
```


```{r}
knn_pred_te <- knn(train = taxi_test[, 1:6],
                 test = taxi_test[, 1:6],
                 cl = taxi_test[, "col"],
                 k = 4
                 )
```

Confusion matrix:
```{r}
confT_te <- table(knn_taxi1, taxi[test_index, "col"])
confT_te
```
Training error:
```{r}
missT <- confT
diag(missT) <- 0 # turn diagonal values to zeros
missCount <- sum(missT) # count the miss-classifications (5)
trainErr_te <- missCount/nrow(taxi_test)
trainErr_te
```










